Tracing Hidden Struggles of Rural Indian Women during Pandemic

### 2024-04-18 outputs/paper.qmd
citations

I obtained the data from the original author to replicate the study. Using methodologies, including fixed-effect analysis and regression models, this paper evaluates the sensitivity of the findings to different analytical approaches. It aims to validate the original results and stimulate further discussion on interpreting data to assess the impacts of public health policies.

This critical analysis aims to clarify the complex relationships between public health funding, policy effectiveness, and their socio-economic effects. By reexamining the original study’s data and methodology, this paper contributes to understanding how public health initiatives can be optimized to serve vulnerable populations during crises.

# Data
## Tools
My replication employed R [@r] along with several essential packages for data manipulation and visualization. Specifically, I utilized Haven [@rHaven] for importing and exporting data, Dplyr [@rDplyr] for data manipulation, Ggplot2 [@rGgplot2] for creating visualizations, Readr [@rReadr] for reading rectangular text data, Here [@rHere] for managing file paths, Janitor [@rJanitor] for examining and cleaning data, KableExtra [@rKableExtra] for constructing complex tables, Knitr [@rKnitr] for dynamic report generation, and Tidyverse [@rTidyverse] for a cohesive data science workflow. Additionally, I employed Coefplot [@rCoefplot] for visualizing regression coefficients and Broom [@rBroom], Lmtest [@rLmtest], and Sandwich [@rSandwich] for regression diagnostics and robust standard errors. For advanced modeling techniques, I utilized Glmnet [@rGlmnet] for Lasso regression.

@@

```{r}
#| label: fig-1
#| fig-cap: "Impact of general economic disruptions on income and women's welfare."
#| echo: false
#| layout-ncol: 2
#| fig-subcap: ["Distributions of Transformed Income", "Effect of COVID-19 on Various Outcomes"]

@@

>>>>

I obtained the data from the original author to replicate the study. Using methodologies, including fixed-effect analysis and regression models, this paper evaluates the sensitivity of the findings to different analytical approaches. This critical analysis aims to clarify the complex relationships between public health funding, policy effectiveness, and their socio-economic effects. By re-examining the original study’s data and methodology, this paper contributes to understanding how public health initiatives can be optimized to serve vulnerable populations during crises.

# Data
## Tools
My replication employed R [@r] along with several essential packages for data manipulation and visualization. Specifically, I utilized Haven [@haven] for importing and exporting data, Dplyr [@dplyr] for data manipulation, Ggplot2 [@ggplot2] for creating visualizations, Readr [@readr] for reading rectangular text data, Here [@here] for managing file paths, Janitor [@rJanitor] for examining and cleaning data, KableExtra [@kableExtra] for constructing complex tables, Knitr [@knitr] for dynamic report generation, and Tidyverse [@tidyverse] for a cohesive data science workflow. Additionally, I employed Coefplot [@rCoefplot] for visualizing regression coefficients and Broom [@broom], Lmtest [@lmtest], and Sandwich [@sandwich] for regression diagnostics and robust standard errors. For advanced modeling techniques, I utilized Glmnet [@glmnet] for Lasso regression.

@@

```{r}
#| label: fig-1
#| fig-cap: "Impact of general economic disruptions on income and women's welfare."
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| fig-subcap: ["Distributions of Transformed Income", "Effect of COVID-19 on Various Outcomes"]

@@

```{r}
#| label: fig-3
#| fig-cap: "Relationship between Containment and Female Well-being or Socioeconomic and Nutritional Outcomes."
#| echo: false
#| warning: false
#| message: false
#| fig-subcap: ["Effect Sizes of Predictors in Mental Health Models", "Effect Sizes of Predictors in Economic and Nutritional Impact Models"]

# Set the data path and read the data
data_path <- "/cloud/project/data/cleaned/"
data_file <- paste0(data_path, "Filtered_Covid_Gender_Data.csv")
data <- read.csv(data_file)

data$demo_caste <- as.character(data$demo_caste)

# Prepare data: factor conversion and recoding
prepare_data <- function(data) {
  data %>%
    mutate(across(c(geo_state, color_zone, demo_caste, demo_religion, fem_marital_status), as.factor),
           demo_caste = factor(recode(as.character(demo_caste), '99' = NA)),
           demo_religion = factor(recode(as.character(demo_religion), '99' = NA)),
           fem_marital_status = factor(recode(as.character(fem_marital_status), '99' = NA)))
}

data <- prepare_data(data)

# Filter data to exclude missing data on necessary variables
filter_data <- function(data, variables) {
  data %>%
    filter(complete.cases(data[, variables])) %>%
    filter(final_status %in% c("1", "2"))
}

# Define general predictors and response variables
general_predictors <- c("cases_per_100000", "deaths_per_100000", "dist_prop_covid_zone", 
                        "red_zone", "orange_zone", "tran_inc_normal", "asset_index")
response_vars <- c("ind_fem_depression_change", "ind_fem_tired_change",
                   "ind_fem_worried_change", "ind_fem_safety_change", "mental_index_change")

data2t <- filter_data(data, c(response_vars, general_predictors, "geo_district"))

rename_vars <- c(
  ind_fem_depression_change = "More Depressed",
  ind_fem_tired_change = "More Tired",
  ind_fem_worried_change = "More Worried",
  ind_fem_safety_change = "Feel Less Safe",
  mental_index_change = "Mental Health Index Change",
  hh_num_inc_reduced = "Household Income Reduced",
  ind_meals_reduced = "Meals Reduced",
  newf_index = "New Financial Index",
  newm_index = "New Mental Health Index",
  cases_per_100000 = "Cases per 100,000",
  deaths_per_100000 = "Deaths per 100,000",
  dist_prop_covid_zone = "District Proportion in Covid Zone",
  red_zone = "Red Zone",
  orange_zone = "Orange Zone",
  tran_inc_normal = "Transition Income Normal",
  asset_index = "Asset Index",
  hh_num_inc_reduced_with_lasso = "Household Income Reduced with LASSO",
  hh_num_inc_reduced_without_lasso = "Household Income Reduced without LASSO",
  ind_meals_reduced_with_lasso = "Meals Reduced with LASSO",
  ind_meals_reduced_without_lasso = "Meals Reduced without LASSO",
  newf_index_with_lasso = "New Financial Index with LASSO",
  newf_index_without_lasso = "New Financial Index without LASSO",
  newm_index_with_lasso = "New Mental Health Index with LASSO",
  newm_index_without_lasso = "New Mental Health Index without LASSO"
)

# Use the updated map in the summarize_models function
summarize_models <- function(models) {
  bind_rows(
    lapply(names(models), function(model_name) {
      model_data <- tidy(models[[model_name]])
      model_data$model <- rename_vars[model_name] %>% coalesce(model_name)  # Rename the model name
      model_data$term <- rename_vars[model_data$term] %>% coalesce(model_data$term)  # Rename the terms
      return(model_data)
    }),
    .id = "model_id"
  ) %>%
    filter(term != "(Intercept)")
}

# Model fitting functions
fit_models <- function(data, response_vars, predictors) {
  models <- list()
  for (response in response_vars) {
    formula <- as.formula(paste(response, "~", paste(predictors, collapse = " + ")))
    model <- glm(formula, family = binomial(), data = data)
    models[[response]] <- model
  }
  return(models)
}

# Fit models for first set of variables
models_response <- fit_models(data2t, response_vars, general_predictors)

# Second set of variables of interest
variables_of_interest <- c("hh_num_inc_reduced", "ind_meals_reduced", "newf_index", "newm_index")
data2t_vars_interest <- filter_data(data, c(variables_of_interest, general_predictors, "geo_district"))

fit_felm_models <- function(data, variables_of_interest, lasso_vars, other_vars) {
  results <- list()
  for (var in variables_of_interest) {
    formula1 <- as.formula(paste0(var, " ~ ", paste(other_vars, collapse = "+"), " | 0 | 0 | geo_district"))
    formula2 <- as.formula(paste0(var, " ~ ", paste(c(other_vars, lasso_vars), collapse = "+"), "| 0 | 0 | geo_district"))
    model1 <- felm(formula1, data = data)
    model2 <- felm(formula2, data = data)
    results[[paste0(var, "_without_lasso")]] <- model1
    results[[paste0(var, "_with_lasso")]] <- model2
  }
  return(results)
}

# Fit models for second set of variables
results_vars_interest <- fit_felm_models(data2t_vars_interest, variables_of_interest, general_predictors, c("dist_prop_covid_zone", "red_zone", "orange_zone"))

# Plotting functions
plot_coefficients_logistic <- function(results) {
  ggplot(results, aes(x = term, y = estimate, color = model)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2, position = position_dodge(width = 0.5)) +
    coord_flip() +
    labs(x = "Predictors", y = "Estimated Coefficients") +
    theme_minimal() +
    theme(legend.title = element_blank(),
          legend.position = "bottom")
}

plot_coefficients_felm <- function(results) {
  ggplot(results, aes(x = term, y = estimate, color = model)) +
    geom_point(position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2, position = position_dodge(width = 0.5)) +
    coord_flip() +
    labs(x = "Predictors", y = "Estimated Coefficients") +
    theme_minimal() +
    theme(legend.title = element_blank(),
          legend.position = "bottom")
}

# Generate and display plots
results_logistic <- summarize_models(models_response)
plot_logistic <- plot_coefficients_logistic(results_logistic)
print(plot_logistic)

results_felm <- summarize_models(results_vars_interest)
plot_felm <- plot_coefficients_felm(results_felm)
print(plot_felm)
```
